---
title: "Reproducible documents"
author: "Naja"
format: html
---

```{r}
2 + 2
```

## About me
- Naja
- Neurologisk forskning
- Aarhus Universitet
Forsker i om PN er **tidligt** symptom på *PD*. Maybe^hvemved^ ingen~endnu~.

## Simple code
```{r}
3 * 3
```

## Testing for Git
Solen skinner og himlen er blå 
```{r setup}
#| message: false
#| warning: false
library(tidyverse)
post_meal_data <- read_csv2(here::here("data/post-meal-insulin.csv"))
small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)
tidier_post_meal_data <- small_post_meal_data |>
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  ) |>
  filter(glykemi == 0) |>
  select(-glykemi)
tidier_post_meal_data
```

## Showing the data
```{r}
post_meal_data
```

glimpse(post_meal_data)

## Plot one continuous variable
```{r}
#| fig-cap: "Distribution of BMI."
#| label: fig-bmi-histo
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```
```

```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
```

## Exercise plot a discrete variable

```{r}
#| fig-cap: "distribution of Group"
#| label: fig-group-bar
ggplot(post_meal_data, aes(x = Group)) +
  geom_bar()
```

See @fig-group-bar above for a cool plot!

## Plotting two discrete variables

```{r}
ggplot(post_meal_data, aes(x = Group, fill = glykemi)) +
  geom_bar()
```

```{r}
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar()
```

## Position af fill ændres til at stå ved siden af frem for oveni, så der bliver 4 søjler
```{r}
ggplot(post_meal_data, aes(x = Group, fill = as.character(glykemi))) +
  geom_bar(position = position_dodge())
```

## Side by side plots
```{r}
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()
ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

```{r}
#| label: fig-bmi-glycemia
#| fig-cap: "BMI and glycemia, side by side."
#| fig-subcap:
#|   - "Distribution of BMI."
#|   - "Number of those with glycemia."
#| layout-ncol: 2
ggplot(post_meal_data, aes(x = BMI)) +
  geom_histogram()

ggplot(post_meal_data, aes(x = glykemi)) +
  geom_bar()
```

## Plotting 2 continuous variables
```{r}
#| fig-cap: "Scatterplot of BMI and the area under the curve for glucose."
#| label: fig-bmi-auc-pg
ggplot(post_meal_data, aes(x = BMI, y = auc_pg)) +
  geom_point() +
  geom_smooth()
```

## Selecting data.
```{r}
select(post_meal_data, Age)
select(post_meal_data, Age, BMI, Weight)
select(post_meal_data, -Group)
```

## Select helpers
```{r}
select(post_meal_data, starts_with("PG"))
select(post_meal_data, ends_with("DXA"))
# All columns containing the number "0"
select(post_meal_data, contains("0"))
small_post_meal_data <- select(post_meal_data, id, Age, BMI, Group, auc_pg, auc_ins, glykemi)
```

## Renaming columns
```{r}
small_post_meal_data |>
  rename(auc_glucose = auc_pg)
small_post_meal_data |>
  rename(
    auc_glucose = auc_pg,
    auc_insulin = auc_ins
  )
```

## Select and rename tasks
```{r}
post_meal_data |>
  select(contains("OGTT"))
post_meal_data |>
  select(contains(".."), -contains("..."))
post_meal_data |>
  rename(
    P.Glucose.minus5.OGTT = P.Glucose..5.OGTT,
    Insulin.minus5.OGTT.X = Insulin..5.OGTT.X
  )
post_meal_data |>
  select(id, BMI, Length, Weight, Age) |>
  rename(Height = Length)
```

## Filtering data
```{r}
small_post_meal_data |>
  filter(glykemi == 1)
small_post_meal_data |>
  filter(BMI >= 25)
small_post_meal_data |>
  filter(Group == "CTR")
small_post_meal_data |>
  filter(Group == "CTR", BMI >= 25)
small_post_meal_data |>
  filter(Group == "CTR" | BMI >= 25)
```
